Quantum H theorem, variational entropy, and relaxation processes

### 2020-12-15 notasparaarticulo.tex
The order of the sections were replaced and some modification were applied

\section*{Abstract}
We review in detail the quantum formulation of the Boltzamnn's H theorem. Starting from a H functional defined in the energy space. By means a variational procedure, we demonstrate the analogous of the quantum H theorem. At difference of the Tolman's procedure to demonstrate the H theorem,  we start from the hypothesis of an out of equilibrium, spatially non homogeneous system, and show that when $\frac{dH}{dt}=0$, the system reaches the equilibrium condition and then proposed H functional becomes the well known expression of the entropy of an ideal quantum gas. The time evolution of the system happens in such a way that $\frac{dH}{dt}<0$. Based on this, we propose a theoretical scheme that allows to describe the time evolution toward the equilibrium of an ideal quantum gas, it is the relaxation process of an out equilibrium, Finally, in the limit of high energy levels, the Boltzmann's $H$ functional is recovered.

@@

\textcolor{red}{The first section of this work consists of obtaining some classical results from Statistical mechanics (the equilibrium distribution function, the Boltzmann transport equation, and an alternative proof of the classical $H$ theorem). The second section consists of obtaining the same results from quantum mechanics. The third consists of analyzing the corresponding principle between the \textit{variational entropy} and Boltzmann's $H$ functional.}

%It is well known for the Physics community that the set of schemes, methods, techniques, and approximations of Statistical Physics to deal with equilibrium systems, are relatively well established. Liouvilles' theorem, Ergodic hypothesis and the main corollary of the H theorem, are some of the conceptual basis on which the theoretical structures of Statistical Thermodynamics have been grown and are sustained \cite{huang, reif, patrick}. This is not the case of the Statistical Physics of out of equilibrium systems, in which Poincares' theorem, Ergodic hypothesis, and H theorem establish the general conceptual basis of the theoretical developments, but these later are not as general and satisfactory as those for equilibrium [ref]. There are of course some well stablished theoretical schemes to deal with out of equilibrium systems like those developed by Onsager known as linear thermodynamics, and some few other restricted developments proposed by some outstanding theoretical authors \cite{kei, onsager}. Nevertheless, even the very fundamentals of out of equilibrium Statistical physics like Boltzmanns' H Theorem, still presents some aspects that deserve some review in order to accomplish better consistency [ref]. In the framework of classical mechanics, the prove of the Boltzmanns' H theorem in the scheme of Kinetic Theory of dilute gases, is founded on the Boltzmann Transport  Equation, which in turn is supported on the Molecular Chaos Hypothesis. 
%It is commonly accepted that BHT solves in a more complete way the Poicares' paradox and concilies the the theoretical conflict between a time reversible microscopic dynamics and an inherent macroscopic one \cite{paradox1, paradox2}. The quantum version of H Theorem, still currently is under analysis and discussion \cite{quantum1, quantum2, quantum3, quantum4}.

%The behavior of the $H$ functional in times bigger than collisions, will be soft and decreasing. When the change of $H$ is zero, the system will be in equilibrium and the corresponding distribution function will be the Maxwell-Boltzmann distribution\cite{huang, patrick}. \\%

%revisar la redaccion y coherencia de estas ideas% 
%One of those discussions is the propose of Tolman. He established the quantum $H$ theorem \cite{tolman} using the probability transition relation, the random phases hypothesis and a homogeneous distribution function. The lack of the quantum Boltzmann transport equation in analogy to classical statistical mechanics has stopped the development of the quantum transport formulation and the quantum non-equilibrium formulation. Another discission is the prove of the quantum $H$ theorem using n-colissions. This idea completes the $H$ theorem in a classical and quantum scheme \cite{binary}. There are some special conditions in quantum mechanics where the second law of thermodynamics is violated \cite{contradictions}, making that the quantum $H$ theorem is not thoroughly understood.\\
%In transport phenomena, the Boltzmann transport equation is solved using the zero and first order approximation. When one uses the zero order approximation, one finds that the total heat flux in the system is zero, that is a contradiction. However, when one applies the first order approximation, one finds the evolution operator expressed by the left expression from the Boltzmann transport equation. That evolution operator tends to zero when time tends to infinity, this is the equilibrium condition \cite{reif}. This result, that describes inhomogeneties, is not expressed in the $H$ theorem.\\
%In order to contribute to a consistent $H$ theorem, in this paper, the quantum and classical $H$ theorem are proved proposing a new $H$ functional for both cases considering a non-homogeneous distribution function and a system divided into imaginary cells. With those assumptions, particles transport is allowed inside the system. Performing variational procedures, one finds that, in the quantum case, Bose-Einstein and Fermi-Dirac distributions minimize the $H$ functional and, in the classical case, Maxwell-Boltzmann distribution minimize the other $H$ functional. A expression of entropy to quantum systems will be obtained in the variational procedure. In other hand, making use of the hypothesis that the system is in a relaxation process and the non-homogeneous distribution function, the $H$ theorem will be proved for both cases.

@@

\textcolor{red}{This section will be divided in three subsections, the first will be dedicated to obtain the equilibrium distribution function (Maxwell-Boltzmann distribution) from the Boltzmann's $H$ functional applying variational procedure.}
\subsection{Maxwell-Boltzmann distribution function}
The Boltzmann's $H$ functional is defined as

@@

    H(t)=\int f(v,t) \ln f(v,t) \mathrm{d}^{3}v,

@@

where $f(v,t)$ is the spatial-homogeneous distribution function of a system. This Boltzmann's $H$ functional describes a system with a global energy $E$, a global number of free

@@

However, we want to consider a $H$ functional with a non-homogeneous distribution function. Because of this, we introduce a modified $H$ functional (denoted as $H'$) defined as

@@

   H'(t)=\sum_{M=1}^{K}\int_{}^{} f_M(v,t) \ln f_M(v,t)\mathrm{d}^3v  \label{CH2}.

@@

This $H'$ describes the same system mention above. Nevertheless, the global system, in turn, consists of a set of $K$ small cells with a constant volume $\delta V_M$, whose local homogeneous distribution functions is $f_{M}(v,t)$. The distribution function will depend on the position of the cells, the velocity $v$, and time $t$. Without loss of generality, we choose these cells to have

@@

$H$ functional, $H_M$, local number of particles, $\mathcal{N}_M$, and local energy, $\mathcal{E}_M$. In terms of the local distribution 
function and the energy $\epsilon(v)$, the last three properties are given by:

@@

    H_M &=&  \int f_M(v,t) \ln f_{M}(v,t) \mathrm{d}^{3}v \label{Hcell},\\
    {\mathcal{N}}_M&=& \int f_{M}(v ,t) \mathrm{d}^{3}v, \nonumber \\
{\mathcal{E}}_M&=& \int f_{M}(v,t)\epsilon(v) \mathrm{d}^{3}v.

@@

@@

    \sum_{M=1}^{K}\int_{}^{}f_M(v,t)\mathrm{d}^3v =N, \ \ \ \sum_{M=1}^{K}\int_{}^{}f_M(v,t)\epsilon(v)\mathrm{d}^3v =E \label{micro}.

@@

    \frac{\delta H'}{\delta f_J(v')}&=&\sum_{M=1}^{K}\int_{}^{}\frac{\delta}{\delta f_J(v')}\left[f_M(v)\ln f_M(v)  \right]\mathrm{d}^3v -\sum_{M=1}^{K}\alpha_M\int_{}^{}\frac{\delta f_M(v)}{\delta f_J(v')}\mathrm{d}^3v \nonumber \\
    &&-\sum_{M=1}^{K}\beta_M \int_{}^{}\epsilon(v)\frac{\delta f_M(v)}{\delta f_J(v')}\mathrm{d}^3v \nonumber \\
    &=&\ln f_J(v')+1-\alpha_J-\beta_J \epsilon(v')=0.

@@

    \ln f(r',v')&=&\alpha(r')+\beta(r') \epsilon(v')-1 \ \ \  \Rightarrow \ \ \ f_J(v')=e^{\alpha_J +\beta_J \epsilon(v')-1} \nonumber \\
    &=&Ce^{\alpha_J+\beta_J \epsilon(v') } \label{relacion1},

@@

As $H'$ in equilibrium \textcolor{red}{is proportional to} entropy, it is trivial to think that Lagrange multipliers do not depend on the position because of the system is in equilibrium and consequently, the distribution function is homogeneous, that is

@@

    f(v)=Ce^{\alpha+\beta \epsilon(v)}\equiv \bar{f}(v).

@@

   H''(t)=\sum_{M=1}^{K}\int_{}^{} \left[f_M(v,t) \ln f_M(v,t)-f_M(v,t)\right]\mathrm{d}^3v  \label{CH3},

@@

  but $\sum_{M=1}^{K} \int f_M(v,t)\mathrm{d}^3v =N$ where $N$ is the total particle number. As $N$ is a constant and any $H$ functional has sense when we calculate its time derivative, this constant can be omitted. \\
  \textcolor{gray}{On the other hand, if the distribution function is spatial-homogeneous, $f_M$ does not depend on the cell number $M$, then we rewrite (\ref{CH2}) as

@@

      H'(t)=\int \sum_{M=1}^{K} [f(v,t)\ln f(v,t)] \mathrm{d}^{3}v = K\int  f(v,t) \ln f(v,t) \mathrm{d}^{3}v= K H_{boltz}(t), 

@@

      H_{boltz}(t)=\int  f(v,t) \ln f(v,t) \mathrm{d}^{3}v.

@@

      \int Kf(v,t) \ln [Kf(v,t)] \mathrm{d}^{3}v = \int [(K\ln K)f(v,t) + Kf(v,t) \ln f(v,t)]\mathrm{d}^{3}v, \label{sum-h}

@@

  we can observe if the number of particles in the $\mu$-space $f(v,t)$ is to big compare with the number of cells $P$, then the first term in (\ref{sum-h}) is negligible, and consequently, we conclude that

@@

      \int f'(v,t) \ln f'(v,t) \mathrm{d}^{3}v = K\int f(v,t) \ln f(v,t) \mathrm{d}^{3}v; \ \ \ f'(v,t)= Kf(v,t). \label{aditive-property} 

@@

  We obtained the additive property when we supposed the distribution function is spatial-homogeneous.}\\
  \textcolor{red}{In the next subsection, we prove the classical $H$ theorem without using the Boltzmann transport equation. It is important to remark that we will use the $H'$ functional. This implies that we are considering a non-homogeneous distribution function.}

%----------------------------------------------

\subsection{Proof of the Boltzmann's $H$ theorem}
\textcolor{red}{To prove the Boltzmann's $H$ theorem (using (\ref{CH2}) as functional), we need to assume that the distribution function $f_M(v,t)$ must satisfy the following assumptions:

@@

        \int f_{M}(v,t) \mathrm{d}^{3}v=\bar{\mathcal{N}}+\Delta_M(t); \ \ \ \ \int f_{M}(v,t) \epsilon(v) \mathrm{d}^{3}v=\bar{\mathcal{E}}+ \delta_M(t), \label{restrictionoutsideclassical}

@@

      \bar{\mathcal{N}}= \int \bar{f}(v) \mathrm{d}^{3}v ; \ \ \ \ \bar{\mathcal{E}}= \int \bar{f}(v)\epsilon(v) \mathrm{d}^{3}v,

@@

  It is important to remark that the quantities $\Delta_M,\delta_M$ are sufficiently big to be different from fluctuations in the system but too small such that the system is not far from the equilibrium state.}  \\

@@

    \frac{dH'}{dt}=\sum_{M=1}^{K}\int_{}^{}\left[ 1+\ln f_M(v,t) \right]\dot f_M(v,t) \mathrm{d}^3v  \label{dH1}.

@@

\textcolor{red}{Using the first-order approximation

@@

   f_{Mn}=\bar{f}_{n}(1+g_{Mn}): \ \ \ \bar{f}_{n}\gg \bar{f}_{n}|g_{Mn}|, \label{firstorder}
\end{eqnarray}}

@@

    \frac{dH'}{dt}=\sum_{M=1}^{K}\int_{}^{}\bar f(v) \left [ 1+\ln \left\{ \bar f(v)+\bar f(v)g_M(v,t) \right\} \right]\dot g_M(v,t)\mathrm{d}^3v  \label{dH1,1}.

@@

\textcolor{red}{We can approximate the logarithm function in (\ref{dH1,1}) to its first-order Taylor series around $\bar f_n g_{nM}=0$ 

@@

    \ln [\bar{f}(v)+\bar{f}(v) g_{M}(v,t)] \approx \ln [\bar{f}(v)]+ g_{M}(v,t). \label{lnapproximationclassical}
\end{equation}}

@@

    \frac{dH'}{dt}&=&\sum_{M=1}^{K} \int_{}^{} \bar f(v)\left[ 1+\ln \bar f(v)+g_M(v,t) \right]\dot g_M(v,t)\mathrm{d}^3v,

@@

    \frac{dH'}{dt}&=&\sum_{M=1}^{K}\int_{}^{}\bar f(v)\left[ \alpha+\beta \epsilon(v) \right]\dot g_M(v,t)\mathrm{d}^3v +\sum_{M=1}^{K}\int_{}^{}\bar f(v)g_M(v,t)\dot g_M(v,t)\mathrm{d}^3v \label{dH1,2}. \nonumber \\
\textcolor{red}{On the other hand, we obtain from the restrictions the following expressions

@@

    \int \bar{f}(v) g_{M}(v,t) \mathrm{d}^{3}v=\Delta_M(t) \ \  &\Rightarrow& \ \  \int \bar{f}(v) \dot{g}_{M}(v,t)\mathrm{d}^{3}v=\dot{\Delta}_M(t), \nonumber \\
    \int  \bar{f}(v) g_{M}(v,t)\epsilon(v) \mathrm{d}^{3}v=\delta_M(t) \ \  &\Rightarrow& \ \  \int \bar{f}(v) \dot{g}_{M}(v,t)\epsilon(v) \mathrm{d}^{3}v=\dot{\delta}_M(t), \nonumber \\

@@

\frac{dH'}{dt}&=&\sum_{M=1}^{K}\int_{}^{}\bar f(v)g_M(v,t)\dot g_M(v,t)\mathrm{d}^3v.

@@

    \frac{dH'}{dt}=\sum_J^{L}\int_{}^{}\bar f(v)g_J^{+}(v,t)\dot g_J^{+}(v,t)\mathrm{d}^3v +\sum_J^{P}\int_{}^{}\bar f(v)g_J^{-}(v,t)\dot g_J^{-}(v,t)\mathrm{d}^3v. \label{classicalH3}

@@

where $L+P=K$. \textcolor{red}{The group of $L$ cells is such that cells have an excess of particles or/and energy (from the mean value in equilibrium). In contrast, the group of $P$ cells is such that cells have to miss particles or/and energy (from the mean value in equilibrium). Besides, $\dot{g}^{+}_{J}$ represents the change on the deviation on cells that have an excess of particles or energy while $\dot{g}^{-}_{J}$  represents the change on the deviations on cells that have missing particles or energy. 

@@

\end{eqnarray}}

@@

                               \sum_J^{L}\int_{}^{}\bar f(v)|g_J^{+}(v,t)||\dot g_J^{+}(v,t)|\mathrm{d}^3v +\sum_J^{P}\int_{}^{}\bar f(v)|g_J^{-}(v,t)||\dot g_J^{-}(v,t)|\mathrm{d}^3v 

@@

We can observe that $\bar{f}(v)$ in (\ref{classicalH4}) is always positive. Also the deviation and its derivative are positive, then all the expression is positive. However, the global sign makes the derivative of the $H'$ with respect to time is always less to zero. With this we proved that $\frac{dH'}{dt}<0$. If the system is in equilibrium, $g(r,v,t)=\dot g(r,v,t)=0$, therefore $\frac{dH'}{dt}=0$.\\

@@

   H'(t)=\sum_{M=1}^{K}\int_{}^{} f_M(v,t) \ln f_M(v,t)\mathrm{d}^3v  \label{CH3},

@@

where $f_M(v,t)$ is the local distribution function of each cell in the system, then in the first-order approximation

@@

This statement will be the classical $H$ theorem with inhomogeneities.
%\textcolor{red}{The last result we want to obtain is the Boltzmann transport equation from the Boltzmann $H$ functional. This procedure is developed in the next subsection.}

%-----------------------------------

\subsection{The Boltzmann transport equation}
The Boltzmann $H$ functional is defined as
\begin{equation}
    H=\int_{}^{}f(\vec{v},t)\ln f(\vec{v},t)\mathrm{d}^{3}v,
\end{equation}
where $f$ is the spatial-homogeneous distribution function of the system and $v$ is the velocity of the particles defined in the $\mu$-space. In order to obtain the Boltzmann transport equation, is necessary to consider that the distribution function is not homogeneous. To obtain the spatial dependence in Boltzmann's transport equation, we need to assume that cells are too small, such that they can be considered as differential volume elements. That is
\begin{equation}
    f_M(\vec{v},t)    \ \ \ \Rightarrow \ \ \ f(\vec{r},\vec{v},t),
\end{equation}
where $r$ is the position of the particles defined in the $\mu$-space. Because of that, we use the following functional
\begin{equation}
    H_{mod}=\int f(\vec{r},\vec{v},t) \ln f(\vec{r},\vec{v},t) \mathrm{d}^{3}v,
\end{equation}
that involves the position dependence.
%\begin{equation}
%    H'=\int_{}^{}\int_{}^{}f(\vec{r},\vec{v},t)\ln f(\vec{r},\vec{v},t)\mathrm{d}^{3}v \mathrm{d}^{3}r.
%\end{equation}
 %This dependence is necessary to obtain the complete Boltzmann transport equation.\\
\\
The following relation must be held in equilibrium 
\begin{equation}
    \delta H_{mod}=0.\label{variation1}
\end{equation}
This due to $H$ functional in equilibrium is proportional to entropy and a maximum.\\ 
Assuming that $\vec{r}=\vec{r}(t)$ and $\vec{v}=\vec{v}(t)$, (\ref{variation1}) casts into
\begin{equation}
    0=\left( 1+\ln f \right)\left( \frac{\partial f}{\partial t}+\dot{\vec{r}}\cdot \nabla f+\dot{\vec{v}}\cdot \nabla_{\vec{v}} f \right).
\end{equation}
The previous relation holds if $(1+\ln f)=0$ or $(\frac{\partial f}{\partial t}+\dot{\vec{r}}\cdot \nabla f+\dot{\vec{v}}\cdot \nabla_{\vec{v}} f)=0$.\\
Assuming that $(1+\ln f)=0$, we find that $f$ is a constant and consequently, $(\frac{\partial f}{\partial t}+\dot{\vec{r}}\cdot \nabla f+\dot{\vec{v}}\cdot \nabla_{\vec{v}} f)=0$. On the other hand, if $f$ is not a constant, the equation of motion will be
\begin{equation}
    \frac{\partial f}{\partial t}+\dot{\vec{r}}\cdot \nabla f+\dot{\vec{v}}\cdot \nabla_{\vec{v}} f=0.\label{transport}
\end{equation}
The equation (\ref{transport}) is hold only in equilibrium. If the system is out of equilibrium, the variation is not equal to zero. As a consequence of the classical $H$ theorem, the variation is less to zero. In conclusion, it is necessary to include a term in (\ref{transport}). This term, as in \cite{huang}, will be the collision term and will be expressed in such form as
\begin{equation}
    \frac{\partial f}{\partial t}+\vec{v}\cdot \nabla f+\frac{\vec{F}}{m}\cdot \nabla_{\vec{v}} f=\left( \frac{df}{dt} \right)_{coll}, \label{classictransport}
\end{equation}
where $\vec{F}$ is an external force, and $m$ is the mass of the particles.\\
\\
The equation (\ref{classictransport}) corresponds to the Boltzmann transport equation. \\
We observed that the variational method and the inclusion of a non-spatial homogeneous distribution function are consistent with the results obtained from Statistical Mechanics. In other words, we proved the classical $H$ theorem and we obtained the Boltzmann transport equation from the Boltzmann $H$ functional.\\
\textcolor{red}{In contrast, there is not quantum analogous of the Boltzmann transport equation. Even more, the analogous quantum $H$ theorem is not clear on the spatial-homogeneous distribution hypothesis. To obtain those results in the quantum mechanic scheme, the next section will develop the same procedure used in this section.}

@@

\end{thebibliography}{}
\clearpage

>>>>

\section*{Abstract}
Starting from a functional, $\mathcal{H}$, defined in terms of non-homogeneous distribution functions, an alternative demonstration of the classical and quantum $H$ theorem is performed \textit{via} a variational procedure. This functional describes the time evolution of out of equilibrium classical or quantum gases, and it satisfies the condition $d\mathcal{H}/dt\leq0$. The equality condition is attained when the system reaches the thermodynamic equilibrium, and in this state -$\mathcal{H}$ becomes the entropy of an ideal  gas. The properties of the functionals $\mathcal{H}$ are investigated, including its correct behavior according to the correspondence principle. In addition, the relaxation of quantum gases towards equilibrium is explored, from which a time-evolution equation is obtained. Thus, the novelty of the approach presented here, is the application of the variational method to a functional defined in terms of non-homogeneous distribution functions, in the $\mu$-space and position-energy spaces respectively.

@@

\textcolor{red}{The first section of this work consists of a review of the Boltzmann transport equation, the Boltzmann $H$ theorem, the Maxwell-Boltzmann distribution, and our alternative form to prove the classical $H$ theorem. The second part consists of a review of the proposal from Tolman about the quantum version of the $H$ theorem, the Bose-Einstein and Fermi-Dirac distributions, and our alternative form to prove the quantum version of the $H$ theorem. The third section consists of a classical-quantum correspondence between the quantum $H$ functional and the classical $H$ functional. Finally, the last section consists of a proposal to explore relaxation processes in the quantum case obtaining the analogous quantum Boltzmann transport equation.}

@@

\subsection{The Boltzmann transport equation}
In the classical version, the Boltzmann transport equation is required to show the $H$-theorem. That equation is

@@

    \left( \frac{\partial }{\partial t}+\vec{v}_1 \cdot \nabla_r +\frac{\vec{F}}{m} \cdot \nabla_{v_1} \right)f_1=\int \mathrm{d}\Omega \int \mathrm{d}^{3}v_2\sigma(\Omega)|\vec{v}_1-\vec{v}_2|(f_2'f_1'-f_2f_1), \label{transport}

@@

where $\Omega$ is the solid angle, $\sigma$ is the scattering section, $F$ the external force applied to the system, $f_1, f_2$ are the distribution function of the particles 1 and 2 respectively before the collision, and $f'_1, f'_2$ are the distribution function of the particles 1 and 2 respectively after the collision.\\
\textcolor{red}{On the one hand, dynamics, inhomogeneities, and external forces are described by the left-hand side of the equation (\ref{transport}). On the other hand, the binary collisions among particles are involved in the right-hand side of the previous equation. However, the \textit{molecular chaos hypothesis} was applied to obtain the right-hand side of (\ref{transport}). The \textit{molecular chaos hypothesis} establishes that spatial and velocity coordinates of the particles are not correlated.}
%-------------------------
\subsection{H-theorem and the Maxwell-Boltzmann distribution}
The $H$ theorem can be proved using the $H$ functional defined by Boltzmann 
   H=\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v,\label{hfunctional}
where $f(\vec{v},t)$ is a homogeneous distribution function. This $H$ functional describes a system with a global energy $E$, a global number of free

@@

\textcolor{red}{The $H$ theorem establishes that if $f(\vec{v},t)$ satisfies the Boltzmann transport equation and the molecular chaos hypothesis, then $\frac{dH}{dt}\leq 0$ and the system arrives to the equilibrium state when $\frac{dH}{dt}=0$. The theorem can be proved by applying the time derivative to (\ref{hfunctional}) and using the Boltzmann transport equation (\ref{transport}). As a consequence of this theorem, the distribution function that describes the system in equilibrium corresponds to the Maxwell-Boltzmann distribution. This distribution can be obtained by solving a sufficient condition for $f$ to obtain the equilibrium distribution function $f_0$

@@

    f_0(\vec{v}'_2)f_0(\vec{v}'_1)- f_0(\vec{v}_2)f_0(\vec{v}_1)=0.

@@

This condition can be obtained from the binary collision term of the Boltzmann transport equation (\ref{transport}).\\
The $H$-theorem fixes the time evolution direction of systems in nature, such as the second law of thermodynamics does. \footnote{The molecular hypothesis is fundamental to determine the physical meaning of the $H$ theorem. For more information, see \cite{huang}.}}
%---------------------------
\subsection{Out of equilibrium, non-homogeneous distributions}
The Boltzmann $H$-functional (\ref{hfunctional}) is defined with a spatial homogeneous distribution $f(\vec{v},t)$. However, we want to consider a $H$ functional with a non-homogeneous distribution function in order to generalize the Boltzmann $H$ theorem. Because of this, we introduce a modified $H$ functional (denoted as $H'$) defined as
   H'(t)=\sum_{M=1}^{K}\int_{}^{} f_M(\vec{v},t) \ln f_M(\vec{v},t)\mathrm{d}^3v  \label{CH2}.
This $H'$ describes a system with a global energy $E$, a global number of free
classical particles $N$, global volume $V$, and a temperature $T$. Nevertheless, the global system, in turn, consists of a set of $K$ small cells with a constant volume $\delta V_M$, whose local homogeneous distribution functions is $f_{M}(\vec{v},t)$. The distribution function will depend on the position of the cells, the velocity $\vec{v}$, and time $t$. Without loss of generality, we choose these cells to have

@@

$H$ functional, $H'_M$, local number of particles, $\mathcal{N}'_M$, and local energy, $\mathcal{E}'_M$. In terms of the local distribution 
function and the energy $\epsilon(\vec{v})$, the last three properties are given by:

@@

    H'_M &=&  \int f_M(\vec{v},t) \ln f_{M}(\vec{v},t) \mathrm{d}^{3}v \label{Hcell},\\
    {\mathcal{N}}'_M&=& \int f_{M}(\vec{v} ,t) \mathrm{d}^{3}v, \nonumber \\
{\mathcal{E}}'_M&=& \int f_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v.

@@

   {\mathcal{N}}'_M=\mathcal{N}'\equiv \bar{\mathcal{N}}'; \ \ \ \  {\mathcal{E}}'_M=\mathcal{E}'\equiv \bar{\mathcal{E}}'.

@@

    \sum_{M=1}^{K}\int_{}^{}f_M(\vec{v},t)\mathrm{d}^3v =N, \ \ \ \sum_{M=1}^{K}\int_{}^{}f_M(\vec{v},t)\epsilon(\vec{v})\mathrm{d}^3v =E \label{micro}.

@@

    \frac{\delta H'}{\delta f_J(\vec{v}')}&=&\sum_{M=1}^{K}\int_{}^{}\frac{\delta}{\delta f_J(\vec{v}')}\left[f_M(\vec{v})\ln f_M(\vec{v})  \right]\mathrm{d}^3v -\sum_{M=1}^{K}\alpha_M\int_{}^{}\frac{\delta f_M(\vec{v})}{\delta f_J(\vec{v}')}\mathrm{d}^3v \nonumber \\
    &&-\sum_{M=1}^{K}\beta_M \int_{}^{}\epsilon(\vec{v})\frac{\delta f_M(\vec{v})}{\delta f_J(\vec{v}')}\mathrm{d}^3v \nonumber \\
    &=&\ln f_J(\vec{v}')+1-\alpha_J-\beta_J \epsilon(\vec{v}')=0.

@@

    \ln f_J(\vec{v}')&=&\alpha_J+\beta_J \epsilon(\vec{v}')-1 \ \ \  \Rightarrow \ \ \ f_J(\vec{v}')=e^{\alpha_J +\beta_J \epsilon(\vec{v}')-1} \nonumber \\
    &=&Ce^{\alpha_J+\beta_J \epsilon(\vec{v}') } \label{relacion1},

@@

As $H'$ in equilibrium is proportional to entropy, it is trivial to think that Lagrange multipliers do not depend on the position because of the system is in equilibrium and consequently, the distribution function is homogeneous, that is

@@

    f(\vec{v})=Ce^{\alpha+\beta \epsilon(\vec{v})}\equiv \bar{f}(\vec{v}).

@@

   H''(t)=\sum_{M=1}^{K}\int_{}^{} \left[f_M(\vec{v},t) \ln f_M(\vec{v},t)-f_M(\vec{v},t)\right]\mathrm{d}^3v  \label{CH3},

@@

  but $\sum_{M=1}^{K} \int f_M(\vec{v},t)\mathrm{d}^3v =N$ where $N$ is the total particle number. As $N$ is a constant and any $H$ functional has sense when we calculate its time derivative, this constant can be omitted. \\
  On the other hand, if the distribution function is spatial-homogeneous, $f_M$ does not depend on the cell number $M$, then we rewrite (\ref{CH2}) as

@@

      H'(t)=\int \sum_{M=1}^{K} [f(\vec{v},t)\ln f(\vec{v},t)] \mathrm{d}^{3}v = K\int  f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v= K H_{boltz}(t), 

@@

      H_{boltz}(t)=\int  f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v.

@@

      \int Kf(\vec{v},t) \ln [Kf(\vec{v},t)] \mathrm{d}^{3}v = \int [(K\ln K)f(\vec{v},t) + Kf(\vec{v},t) \ln f(\vec{v},t)]\mathrm{d}^{3}v, \label{sum-h}

@@

  we can observe if the number of particles in the $\mu$-space $f(\vec{v},t)$ is to big compare with the number of cells $P$, then the first term in (\ref{sum-h}) is negligible, and consequently, we conclude that

@@

      \int f'(\vec{v},t) \ln f'(\vec{v},t) \mathrm{d}^{3}v = K\int f(\vec{v},t) \ln f(\vec{v},t) \mathrm{d}^{3}v; \ \ \ f'(\vec{v},t)= Kf(\vec{v},t). \label{aditive-property} 

@@

  We obtained the additive property when we supposed the distribution function is spatial-homogeneous.\\
To prove the Boltzmann's $H$ theorem (using (\ref{CH2}) as functional), we need to assume that the distribution function $f_M(\vec{v},t)$ must satisfy the following assumptions:

@@

        \int f_{M}(\vec{v},t) \mathrm{d}^{3}v=\bar{\mathcal{N}}+\Delta_M(t); \ \ \ \ \int f_{M}(\vec{v},t) \epsilon(\vec{v}) \mathrm{d}^{3}v=\bar{\mathcal{E}}+ \delta_M(t), \label{restrictionoutsideclassical}

@@

      \bar{\mathcal{N}}= \int \bar{f}(\vec{v}) \mathrm{d}^{3}v ; \ \ \ \ \bar{\mathcal{E}}= \int \bar{f}(\vec{v})\epsilon(\vec{v}) \mathrm{d}^{3}v,

@@

  It is important to remark that the quantities $\Delta_M,\delta_M$ are sufficiently big to be different from fluctuations in the system but too small such that the system is not far from the equilibrium state.  \\

@@

    \frac{dH'}{dt}=\sum_{M=1}^{K}\int_{}^{}\left[ 1+\ln f_M(\vec{v},t) \right]\dot f_M(\vec{v},t) \mathrm{d}^3v  \label{dH1}.

@@

Using the first-order approximation

@@

   f_{Mn}(\vec{v},t)=\bar{f}_{n}(\vec{v})(1+g_{Mn}(\vec{v},t)): \ \ \ \bar{f}_{n}(\vec{v})\gg \bar{f}_{n}(\vec{v})|g_{Mn}(\vec{v},t)|, \label{firstorder}

@@

    \frac{dH'}{dt}=\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v}) \left [ 1+\ln \left\{ \bar f(\vec{v})+\bar f(\vec{v})g_M(\vec{v},t) \right\} \right]\dot g_M(\vec{v},t)\mathrm{d}^3v  \label{dH1,1}.

@@

We can approximate the logarithm function in (\ref{dH1,1}) to its first-order Taylor series around $\bar f_n(\vec{v}) g_{nM}(\vec{v},t)=0$ 

@@

    \ln [\bar{f}(\vec{v})+\bar{f}(\vec{v}) g_{M}(\vec{v},t)] \approx \ln [\bar{f}(\vec{v})]+ g_{M}(\vec{v},t). \label{lnapproximationclassical}

@@

    \frac{dH'}{dt}&=&\sum_{M=1}^{K} \int_{}^{} \bar f(\vec{v})\left[ 1+\ln \bar f(\vec{v})+g_M(\vec{v},t) \right]\dot g_M(\vec{v},t)\mathrm{d}^3v,

@@

    \frac{dH'}{dt}&=&\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v})\left[ \alpha+\beta \epsilon(\vec{v}) \right]\dot g_M(\vec{v},t)\mathrm{d}^3v +\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v})g_M(\vec{v},t)\dot g_M(\vec{v},t)\mathrm{d}^3v \label{dH1,2}. \nonumber \\

@@

    \int \bar{f}(\vec{v}) g_{M}(\vec{v},t) \mathrm{d}^{3}v=\Delta_M(t) \ \  &\Rightarrow& \ \  \int \bar{f}(\vec{v}) \dot{g}_{M}(\vec{v},t)\mathrm{d}^{3}v=\dot{\Delta}_M(t), \nonumber \\
    \int  \bar{f}(\vec{v}) g_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v=\delta_M(t) \ \  &\Rightarrow& \ \  \int \bar{f}(\vec{v}) \dot{g}_{M}(\vec{v},t)\epsilon(\vec{v}) \mathrm{d}^{3}v=\dot{\delta}_M(t), \nonumber \\

@@

\frac{dH'}{dt}&=&\sum_{M=1}^{K}\int_{}^{}\bar f(\vec{v})g_M(\vec{v},t)\dot g_M(\vec{v},t)\mathrm{d}^3v.

@@

    \frac{dH'}{dt}=\sum_J^{L}\int_{}^{}\bar f(\vec{v})g_J^{+}(\vec{v},t)\dot g_J^{+}(\vec{v},t)\mathrm{d}^3v +\sum_J^{P}\int_{}^{}\bar f(\vec{v})g_J^{-}(\vec{v},t)\dot g_J^{-}(\vec{v},t)\mathrm{d}^3v. \label{classicalH3}

@@

where $L+P=K$. The group of $L$ cells is such that cells have an excess of particles or/and energy (from the mean value in equilibrium). In contrast, the group of $P$ cells is such that cells have to miss particles or/and energy (from the mean value in equilibrium). Besides, $\dot{g}^{+}_{J}$ represents the change on the deviation on cells that have an excess of particles or energy while $\dot{g}^{-}_{J}$  represents the change on the deviations on cells that have missing particles or energy. 

@@

@@

                               \sum_J^{L}\int_{}^{}\bar f(\vec{v})|g_J^{+}(\vec{v},t)||\dot g_J^{+}(\vec{v},t)|\mathrm{d}^3v +\sum_J^{P}\int_{}^{}\bar f(\vec{v})|g_J^{-}(\vec{v},t)||\dot g_J^{-}(\vec{v},t)|\mathrm{d}^3v 

@@

We can observe that $\bar{f}(\vec{v})$ in (\ref{classicalH4}) is always positive. Also the deviation and its derivative are positive, then all the expression is positive. However, the global sign makes the derivative of the $H'$ with respect to time is always less to zero. With this we proved that $\frac{dH'}{dt}<0$. If the system is in equilibrium, $g_{J}(\vec{v},t)=\dot g_J(\vec{v},t)=0$, therefore $\frac{dH'}{dt}=0$.\\

@@

   H'(t)=\sum_{M=1}^{K}\int_{}^{} f_M(\vec{v},t) \ln f_M(\vec{v},t)\mathrm{d}^3v  \label{CH3},

@@

where $f_M(\vec{v},t)$ is the local distribution function of each cell in the system, then in the first-order approximation

@@

This statement will be the classical $H$ theorem with inhomogeneities.\\
\textcolor{red}{In the next section, we present a proposal quantum version of the $H$ theorem defined by Tolman and our proposal quantum version of the $H$ theorem applying the method of the volume divided into cells.}

@@

\end{thebibliography}{}

\end{document}
